Online appendix for "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences

### 2020-10-08 manuscript/manuscript.Rmd
Final changes for NVSQ

---
title: "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences"
short-title: Dynamics of International Giving
author:
- name: Suparna Chaudhry
  affiliation: "Lewis & Clark College"
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "September 29, 2020"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State restrictions on non-governmental organizations (NGOs) have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private philanthropy. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and main funding sources. We discuss the implications of our results on how nonprofits can use different framing appeals to increase fundraising at a time when traditional public donor funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,012"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
---

@@

Recent experimental evidence also finds little support for the idea that individual donor behavior mimics that of official donors. @desaikharas find that unlike official donors who tend to channel funds to countries with better institutional quality or to reward governmental performance [@Bermeo:2011; @DietrichWright:2015], individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to NGOs in countries experiencing a humanitarian crisis—in particular, crises following natural disasters. @desaikharas conclude that compared to donor agencies, "private donors respond to different project and country characteristics." (p. 517) International philanthropy by individuals should therefore be investigated on its own terms.

@@

We argue that structural factors—including whether an organization works abroad and the domestic political environment of the group's host country—can serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs and their host governments serves one such heuristic. NGO legislation is not inherently restrictive, and governments frequently use laws to regulate the behavior of NGOs [@BloodgoodTB; @DeMattee:2018]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

@@

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like the United States Agency for International Development (USAID) because such funding can signal a lack of independence from the government. Organizations that receive substantial funding from their home governments frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help and that their marginal donation would make a noticeable difference.

@@

Each of the varied treatments correspond to our hypotheses: crackdown vs. no crackdown, humanitarian assistance vs. human rights, and government vs. nongovernmental funding. Participants might misinterpret the vignette and assume that IRC's host governments fund the organization rather than its home government. In either case, knowing about government funding should still serve as a signal of the organization's deservingness. Similarly, there could be confusion with phrase "private funding", but since we use it as the opposite of government funding, it should still act as a heuristic. Free responses tend to confirm the correct interpretation of the vignette, and many respondents explicitly justified their lack of support for IRC because of government funding, as expected ("I see that a large portion of its funding comes from government donors, so I feel it doesn't need me as much."[^6216778])

>>>>

---
title: "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences"
short-title: Dynamics of International Giving
author:
- name: Suparna Chaudhry
  affiliation: "Lewis & Clark College"
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "October 7, 2020"
published: "Forthcoming in *Nonprofit and Voluntary Sector Quarterly* (NVSQ)"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  State restrictions on non-governmental organizations (NGOs) have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private philanthropy. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and main funding sources. We discuss the implications of our results on how nonprofits can use different framing appeals to increase fundraising at a time when traditional public donor funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input, as well as the anonymous reviewers at Nonprofit and Voluntary Sector Quarterly (NVSQ) for their excellent feedback. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,165"
keywords: nonprofit law, NGO crackdown, international philanthropy, charitable giving, survey experiments
---

@@

Recent experimental evidence also finds little support for the idea that individual donor behavior mimics that of official donors. @desaikharas find that unlike official donors who tend to channel funds to countries with better institutional quality or to reward governmental performance [@Bermeo:2011; @DietrichWright:2015], individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to NGOs in countries experiencing a humanitarian crisis—in particular, crises following natural disasters. @desaikharas conclude that compared to donor agencies, "private donors respond to different project and country characteristics" (p. 517). International philanthropy by individuals should therefore be investigated on its own terms.

@@

We argue that structural factors—including whether an organization works abroad and the domestic political environment of the group's host country—can serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs and their host governments serves one such heuristic. NGO legislation is not inherently restrictive, and governments frequently use laws to regulate the behavior of NGOs [@BloodgoodTB; @DeMattee:2018]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy [@Spires:2020]. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

@@

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like the United States Agency for International Development (USAID) because such funding can signal a lack of independence from the government. Organizations that receive substantial funding from their home governments frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012; @YuShenLi:2020]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help and that their marginal donation would make a noticeable difference.

@@

Each of the varied treatments correspond to our hypotheses: crackdown vs. no crackdown, humanitarian assistance vs. human rights, and government vs. nongovernmental funding. Participants might misinterpret the vignette and assume that IRC's host governments fund the organization rather than its home government. In either case, knowing about government funding should still serve as a signal of the organization's deservingness. Similarly, there could be confusion with phrase "private funding," but since we use it as the opposite of government funding, it should still act as a heuristic. Free responses tend to confirm the correct interpretation of the vignette, and many respondents explicitly justified their lack of support for IRC because of government funding, as expected ("I see that a large portion of its funding comes from government donors, so I feel it doesn't need me as much."[^6216778])

### 2020-10-08 manuscript/supplement.Rmd
Final changes for NVSQ

---
title: 'Online appendix for "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences"'
short-title: 'Online appendix for "Dynamics of International Giving"'
author:
- name: Suparna Chaudhry
  affiliation: "Lewis & Clark College"
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
title-page: false
published: 
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: true
appendix: true
---

@@

```{r child = "../analysis/text/experiment_blind.md"}
```

>>>>

---
title: 'Online appendix for "Dynamics of International Giving: How Heuristics Shape Individual Donor Preferences"'
short-title: 'Online appendix for "Dynamics of International Giving"'
author:
- name: Suparna Chaudhry
  affiliation: "Lewis & Clark College"
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "October 7, 2020"
title-page: false
published: 
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: true
appendix: true
---

@@

```{r child = "../analysis/text/experiment.md"}
```
